### 2022-07-21 figures_Paper1/setup.tex

\documentclass{standalone}

\usepackage{amsmath}
\usepackage{pgfplots}
\pgfplotsset{compat=1.16}
\usepackage{tikz-3dplot}
\usetikzlibrary{calc}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{decorations.markings}
\usetikzlibrary{decorations.pathmorphing}

\usepackage{xcolor}
\usepackage{amsmath}

\definecolor{hblue}{RGB}{0,112,255}    % Brandeis blue
\definecolor{hred}{RGB}{199,44,72}     % raspberry red
\definecolor{hgreen}{RGB}{50,205,50}   % lime green
\definecolor{horange}{RGB}{255,127,0}  % orange
\definecolor{hpurple}{RGB}{111,45,168} % grape
\definecolor{hyellow}{RGB}{255,211,0}  % cyber yellow
\definecolor{hgray}{HTML}{323232}      % anthrazit gray

\begin{document}
\begin{tikzpicture}[>=Stealth]
\begin{axis}[
axis lines=none,
view={150}{10},
xmin=-1,xmax=6,
ymin=0,ymax=4,
zmin=-2,zmax=3.,
xlabel=$x$,
ylabel=$y$,
zlabel=$z$,
scale=1
]

% right lead 
% balls

\foreach \x in {-1,0}          % loop for balls 
  \foreach \y in {0,1,...,3}
    \foreach \z in {0,1,...,3}
       \addplot3 [only marks,mark=ball,mark size=2pt,ball color=hblue,line width=0.1pt] coordinates{(\x,\y,\z)};

% lines

% (axis cs:x,y,z)

\pgfplotsinvokeforeach{0,1,...,3} {
   \draw[color=hblue] (axis cs:-1,#1,0) -- (axis cs:-1,#1,3);
   \draw[color=hblue] (axis cs:-1,#1,0) -- (axis cs:0,#1,0);
   \draw[color=hblue] (axis cs:-1,#1,1) -- (axis cs:0,#1,1);
   \draw[color=hblue] (axis cs:-1,#1,2) -- (axis cs:0,#1,2);
   \draw[color=hblue] (axis cs:-1,#1,3) -- (axis cs:0,#1,3);
};

% lines

\pgfplotsinvokeforeach{0,1,...,3} {
   \draw[color=hblue] (axis cs:-1,0,#1) -- (axis cs:-1,3,#1);
};
\pgfplotsinvokeforeach{0,1,...,3} {
   \draw[color=hblue] (axis cs:0,#1,0) -- (axis cs:0,#1,3);
};
\pgfplotsinvokeforeach{0,1,...,3} {
   \draw[color=hblue] (axis cs:0,0,#1) -- (axis cs:0,3,#1);
};

% bottom
% \fill[color=hblue,opacity=0.2] (axis cs:1,0,0) -- (axis cs:0,0,0) -- (axis cs:0,3,0) -- (axis cs:1,3,0) -- cycle;
% back
% \fill[color=hblue,opacity=0.2] (axis cs:1,0,0) -- (axis cs:0,0,0) -- (axis cs:0,0,3) -- (axis cs:1,0,3) -- cycle;

% front
\fill[color=hblue,opacity=0.5] (axis cs:0,0,0) -- (axis cs:0,3,0) -- (axis cs:0,3,3) -- (axis cs:0,0,3) -- cycle;
% left side
\fill[color=hblue,opacity=0.5] (axis cs:0,3,0) -- (axis cs:-1,3,0) -- (axis cs:-1,3,3) -- (axis cs:0,3,3) -- cycle;
% top
\fill[color=hblue,opacity=0.5] (axis cs:0,3,3) -- (axis cs:-1,3,3) -- (axis cs:-1,0,3) -- (axis cs:0,0,3) -- cycle;

% interface right bath

\pgfplotsinvokeforeach{0,1,...,3} {
   \draw[color=horange] (axis cs:0,#1,0) -- (axis cs:1,#1,0);
   \draw[color=horange] (axis cs:0,#1,1) -- (axis cs:1,#1,1);
   \draw[color=horange] (axis cs:0,#1,2) -- (axis cs:1,#1,2);
   \draw[color=horange] (axis cs:0,#1,3) -- (axis cs:1,#1,3);
};

% ni-layer belonging to right bath

\fill[color=hred,opacity=0.5] (axis cs:1,0,0) -- (axis cs:1,3,0) -- (axis cs:1,3,3) -- (axis cs:1,0,3) -- cycle;

% lines
\pgfplotsinvokeforeach{0,1,...,3} {
   \draw[color=hred] (axis cs:1,#1,0) -- (axis cs:1,#1,3);
};
\pgfplotsinvokeforeach{0,1,...,3} {
   \draw[color=hred] (axis cs:1,0,#1) -- (axis cs:1,3,#1);
};

% balls
\foreach \y in {0,1,...,3}
  \foreach \z in {0,1,...,3}
     \addplot3 [only marks,mark=ball,mark size=2pt,ball color=hred,line width=0.1pt] coordinates{(1,\y,\z)};

% interface ni layer - Hubbard layer
\pgfplotsinvokeforeach{0,1,...,3} {
   \draw[color=black] (axis cs:1,#1,0) -- (axis cs:2,#1,0);
   \draw[color=black] (axis cs:1,#1,1) -- (axis cs:2,#1,1);
   \draw[color=black] (axis cs:1,#1,2) -- (axis cs:2,#1,2);
   \draw[color=black] (axis cs:1,#1,3) -- (axis cs:2,#1,3);
};

% central layer
\fill[color=hgreen,opacity=0.5] (axis cs:2,0,0) -- (axis cs:2,3,0) -- (axis cs:2,3,3) -- (axis cs:2,0,3) -- cycle;
% lines
\pgfplotsinvokeforeach{0,1,...,3} {
   \draw[color=hgreen] (axis cs:2,#1,0) -- (axis cs:2,#1,3);
};
\pgfplotsinvokeforeach{0,1,...,3} {
   \draw[color=hgreen] (axis cs:2,0,#1) -- (axis cs:2,3,#1);
};
% balls
\foreach \y in {0,1,...,3}
  \foreach \z in {0,1,...,3}
     \addplot3 [only marks,mark=ball,mark size=3pt,ball color=hgreen,line width=0.1pt] coordinates{(2,\y,\z)};

% interface ni layer - Hubbard layer
\pgfplotsinvokeforeach{0,1,...,3} {
   \draw[color=black] (axis cs:2,#1,0) -- (axis cs:3,#1,0);
   \draw[color=black] (axis cs:2,#1,1) -- (axis cs:3,#1,1);
   \draw[color=black] (axis cs:2,#1,2) -- (axis cs:3,#1,2);
   \draw[color=black] (axis cs:2,#1,3) -- (axis cs:3,#1,3);
};

% ni-layer belonging to left bath

\fill[color=hred,opacity=0.5] (axis cs:3,0,0) -- (axis cs:3,3,0) -- (axis cs:3,3,3) -- (axis cs:3,0,3) -- cycle;

% lines
\pgfplotsinvokeforeach{0,1,...,3} {
   \draw[color=hred] (axis cs:3,#1,0) -- (axis cs:3,#1,3);
};
\pgfplotsinvokeforeach{0,1,...,3} {
   \draw[color=hred] (axis cs:3,0,#1) -- (axis cs:3,3,#1);
};

% balls
\foreach \y in {0,1,...,3}
  \foreach \z in {0,1,...,3}
     \addplot3 [only marks,mark=ball,mark size=2pt,ball color=hred,line width=0.1pt] coordinates{(3,\y,\z)};

% interface left bath

\pgfplotsinvokeforeach{0,1,...,3} {
   \draw[color=horange] (axis cs:3,#1,0) -- (axis cs:4,#1,0);
   \draw[color=horange] (axis cs:3,#1,1) -- (axis cs:4,#1,1);
   \draw[color=horange] (axis cs:3,#1,2) -- (axis cs:4,#1,2);
   \draw[color=horange] (axis cs:3,#1,3) -- (axis cs:4,#1,3);
};

% left lead
% balls
\foreach \x in {4,5}
  \foreach \y in {0,1,...,3}
    \foreach \z in {0,1,...,3}
       \addplot3 [only marks,mark=ball,mark size=2pt,ball color=hblue,line width=0.1pt] coordinates{(\x,\y,\z)};

% lines
\pgfplotsinvokeforeach{0,1,...,3} {
   \draw[color=hblue] (axis cs:4,#1,0) -- (axis cs:4,#1,3);
   \draw[color=hblue] (axis cs:4,#1,0) -- (axis cs:5,#1,0);
   \draw[color=hblue] (axis cs:4,#1,1) -- (axis cs:5,#1,1);
   \draw[color=hblue] (axis cs:4,#1,2) -- (axis cs:5,#1,2);
   \draw[color=hblue] (axis cs:4,#1,3) -- (axis cs:5,#1,3);
};

\pgfplotsinvokeforeach{0,1,...,3} { % z-direction
   \draw[color=hblue] (axis cs:4,0,#1) -- (axis cs:4,3,#1);
};
\pgfplotsinvokeforeach{0,1,...,3} {
   \draw[color=hblue] (axis cs:5,#1,0) -- (axis cs:5,#1,3);
};
\pgfplotsinvokeforeach{0,1,...,3} {
   \draw[color=hblue] (axis cs:5,0,#1) -- (axis cs:5,3,#1);
};

% bottom
% \fill[color=hblue,opacity=0.5] (axis cs:4,0,0) -- (axis cs:3,0,0) -- (axis cs:3,3,0) -- (axis cs:4,3,0) -- cycle;
% back
% \fill[color=hblue,opacity=0.5] (axis cs:4,0,0) -- (axis cs:3,0,0) -- (axis cs:3,0,3) -- (axis cs:4,0,3) -- cycle;

% front
\fill[color=hblue,opacity=0.5] (axis cs:5,0,0) -- (axis cs:5,3,0) -- (axis cs:5,3,3) -- (axis cs:5,0,3) -- cycle;
% left side
\fill[color=hblue,opacity=0.5] (axis cs:5,3,0) -- (axis cs:4,3,0) -- (axis cs:4,3,3) -- (axis cs:5,3,3) -- cycle;
% top
\fill[color=hblue,opacity=0.5] (axis cs:5,3,3) -- (axis cs:4,3,3) -- (axis cs:4,0,3) -- (axis cs:5,0,3) -- cycle;

% coordinates for light

\coordinate (top1) at (axis cs:2,0.5,3);
\coordinate (top2) at (axis cs:2,1.5,3);
\coordinate (top3) at (axis cs:2,2.5,3);

% arrows for hoppings

\node at (axis cs:-0.1,3.5,-0.15) {\textcolor{hblue}{$\mu_r$}};
\node at (axis cs:5,3.5,-0.15) {\textcolor{hblue}{$\mu_l$}};
\node at (axis cs:2,2.8,-0.35) {\textcolor{hgreen}{$U,g$}};
\node at (axis cs:3.5,3,-0.25) {\textcolor{horange}{$\Gamma$}};
\node at (axis cs:1.3,3,-0.25) {\textcolor{black}{$v$}};

\end{axis}

% cw-light

\draw [->,>=Stealth,line width=2pt,decorate,decoration={snake,amplitude=0.2cm,segment length=0.4cm,pre length=0.4cm,post length=0.4cm},draw=hyellow] ($ (top1)+(120:2) $) -- (top1);
\draw [->,>=Stealth,line width=2pt,decorate,decoration={snake,amplitude=0.2cm,segment length=0.4cm,pre length=0.4cm,post length=0.4cm},draw=hyellow] ($ (top2)+(115:2) $) -- (top2);
\draw [->,>=Stealth,line width=2pt,decorate,decoration={snake,amplitude=0.2cm,segment length=0.4cm,pre length=0.4cm,post length=0.4cm},draw=hyellow] ($ (top3)+(110:2) $) -- (top3);
% \draw [->,>=Stealth,line width=2pt,decorate,decoration={snake,amplitude=0.2cm,segment length=0.8cm,pre length=0.4cm,post length=0.4cm},draw=hyellow] ($ (top4)+(105:3) $) -- (top4);

% pulsed light
% \begin{scope}[decoration={markings,
% mark=at position 0.0 with {{\arrowreversed[hyellow,xshift=-6]{Stealth}}}},
% domain=-1.6:1.5,samples=300,xshift=8.4cm,yshift=4.6cm,rotate=110]
% \draw[postaction={decorate},draw=hyellow,line width=2pt] plot (\x,{0.8*exp(-\x*\x/(2.*0.3))*sin(10.*\x r)});
% \end{scope}

\end{tikzpicture}
\end{document}