### 2022-07-21 figures_Paper1/spec_filling_mu1_v_0.4_0.8_O_11_eph.tex

\documentclass[10pt]{standalone}
\usepackage[english]{babel} % language
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}

%%%%%%%%%%%%
% PGFPLOTS %
%%%%%%%%%%%%
\usepackage{pgfplots}    % pgfplots 
\usepackage{tikz-3dplot} % 3D coordinate frames
\usepackage{gnuplot-lua-tikz}
\usepgfplotslibrary{fillbetween}
\usepgfplotslibrary{colormaps}
\pgfplotsset{compat=1.17}
% \pgfplotsset{trig format plots=rad}

%%%%%%%%%%%%%%%%% 
% TikZ packages %
%%%%%%%%%%%%%%%%%
\usetikzlibrary{math}
\usetikzlibrary{calc}
\usetikzlibrary{decorations.markings}
\usetikzlibrary{decorations.pathmorphing}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{positioning}
\usetikzlibrary{fit}
\usetikzlibrary{matrix}
\usetikzlibrary{intersections}
\usetikzlibrary{bending}
\usetikzlibrary{pgfplots.groupplots}

\usepackage{xcolor}
\usepackage{braket}

% alternative
%\definecolor{hturquoise}{RGB}{0,206,209}   %turquoise
%\definecolor{hred}{RGB}{255,0,0}             % red 
%\definecolor{hbrown}{RGB}{128,70,27}    % russet brown 
%\definecolor{hpurple}{RGB}{111,45,168} % grape
%\definecolor{hyellow}{RGB}{255,255,0}   % yellow
\definecolor{hgray}{HTML}{323232}      % anthrazit gray 
%\definecolor{hgreen}{RGB}{50,205,50}   % dark lime green 
%\definecolor{hmagenta}{RGB}{255,0,255}   % magenta

\definecolor{hcoral}{RGB}{255,127,80}     % coral
\definecolor{hblue}{RGB}{0,112,255}        % Brandeis blue
\definecolor{hred}{RGB}{199,44,72}         % raspberry red
\definecolor{hlime}{RGB}{0,255,0}            % lime green
\definecolor{hyellow}{RGB}{255,211,0}     % cyber yellow
\definecolor{hcyan}{RGB}{0,255,255}       % bright cyan

\DeclareMathAlphabet{\pazocal}{OMS}{zplm}{m}{n}
\newcommand{\capcali}{\pazocal{I}}
\newcommand{\capcalr}{\pazocal{R}}

\begin{document}

\begin{tikzpicture}[>=Stealth]

% foreground, background
\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfdeclarelayer{main} 
\pgfsetlayers{background,main,foreground}

\begin{groupplot}[
group style={
group size=2 by 1,
xticklabels at=edge bottom,  
yticklabels at=edge left,
%vertical sep=10pt,
horizontal sep=20pt
}, 
minor tick num=1, 
width=0.6\textwidth,
height=0.6\textwidth    
]

% 1° figure

\nextgroupplot[
xmin=-15.,      
xmax=15., 
ymin=0.00, 
ymax=0.09,   
minor tick num=1, 
% grid=both,
xlabel={$\omega/t^{*}$},
ylabel={$A(\omega)/t^{*-1}$},
legend style={at={(axis cs:-9.3,0.017)},anchor=north,mark options={scale=0.85},/tikz/every odd column/.append style={column sep=0.5mm},legend columns=1},
legend cell align=left,
clip mode=individual,
clip marker paths=true, 
xtick={-15,-10,-5,0,5,10,15},
ytick={0.00,0.03,0.06,0.09},
% yticklabels={\smash{\raisebox{0.1\height}{0}},0.002,0.004,0.006,0.008,0.01,0.012},
% yticklabels={\smash{\raisebox{0.1\height}{0}},0.002,0.004,0.006,0.008,0.01,0.012,\smash{\raisebox{-0.8\height}{5}}},
x tick label style={
    /pgf/number format/.cd,
    fixed,
    fixed zerofill,
    precision=0,
    /tikz/.cd},
y tick label style={
    /pgf/number format/.cd,
    fixed,
    fixed zerofill, 
    precision=0, 
    /tikz/.cd}
]

\addlegendimage{solid,color=hblue,scale={1.0},line width=1pt}
\addlegendentry{$el$};
\addlegendimage{solid,color=hred,scale={1.0},line width=1pt} 
\addlegendentry{$el+ph$};

% % % % % % % %  PLOTS

% \draw (axis cs:4,0) -- (axis cs:10,0.01);

\addplot[solid,raw gnuplot,color=hblue,line width=1pt] gnuplot {
plot './Green_hyb0.4O11.dat' u 1:(-1/3.1415)*($3) w l;
};

\addplot[dashed,raw gnuplot,color=hblue,line width=1pt] gnuplot {
plot './neq_distr_f_filling_hyb0.4O11.dat' u 1:3 w l;
};

\addplot[solid,raw gnuplot,color=hred,line width=1pt] gnuplot {
plot './Green_hyb0.4O11ph.dat' u 1:(-1/3.1415)*($3) w l;
};

\addplot[dashed,raw gnuplot,color=hred,line width=1pt] gnuplot {
plot './neq_distr_f_filling_hyb0.4O11ph.dat' u 1:3 w l;
};

 \node[fill=none,scale=1.0] at (axis cs:-13.5,0.085) {\bfseries(a)};
 \node[fill=none,scale=1.0] at (axis cs:11.5,0.085) {$v=0.4$};

% 2° figure

\nextgroupplot[
xmin=-15.,     
xmax=15., 
ymin=0.00, 
ymax=0.09,   
minor tick num=1, 
% grid=both,
xlabel={$\omega/t^{*}$},
% ylabel={$A(\omega)/t^{*-1}$},
legend style={at={(axis cs:-9.3,0.017)},anchor=north,mark options={scale=0.85},/tikz/every odd column/.append style={column sep=0.5mm},legend columns=1},
legend cell align=left,
clip mode=individual,
clip marker paths=true, 
xtick={-15,-10,-5,0,5,10,15},
ytick={0.00,0.03,0.06,0.09},
% yticklabels={\smash{\raisebox{0.1\height}{0}},0.002,0.004,0.006,0.008,0.01,0.012},
% yticklabels={\smash{\raisebox{0.1\height}{0}},0.002,0.004,0.006,0.008,0.01,0.012,\smash{\raisebox{-0.8\height}{5}}},
x tick label style={
    /pgf/number format/.cd,
    fixed,
    fixed zerofill,
    precision=0,
    /tikz/.cd},
y tick label style={
    /pgf/number format/.cd,
    fixed,
    fixed zerofill,
    precision=0,
    /tikz/.cd}
]

\addlegendimage{solid,color=hcoral,scale={1.0},line width=1pt}
\addlegendentry{$el$};
\addlegendimage{solid,color=hlime,scale={1.0},line width=1pt} 
\addlegendentry{$el+ph$};

% % % % % % % %  PLOTS

% \draw (axis cs:4,0) -- (axis cs:10,0.01);

\addplot[solid,raw gnuplot,color=hcoral,line width=1pt] gnuplot {
plot './Green_hyb0.8O11.dat' u 1:(-1/3.1415)*($3) w l;
};

\addplot[dashed,raw gnuplot,color=hcoral,line width=1pt] gnuplot {
plot './neq_distr_f_filling_hyb0.8O11.dat' u 1:3 w l;
};

\addplot[solid,raw gnuplot,color=hlime,line width=1pt] gnuplot {
plot './Green_hyb0.8O11ph.dat' u 1:(-1/3.1415)*($3) w l;
};

\addplot[dashed,raw gnuplot,color=hlime,line width=1pt] gnuplot {
plot './neq_distr_f_filling_hyb0.8O11ph.dat' u 1:3 w l;
};

 \node[fill=none,scale=1.0] at (axis cs:-13.5,0.085) {\bfseries(b)};
 \node[fill=none,scale=1.0] at (axis cs:11.5,0.085) {$v=0.8$};

\end{groupplot}

 \end{tikzpicture}

\end{document}